---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

So, to start with you need to lod the data and turn dates and intervals into as.date and as.factor.
```{r}
activity = read.csv("activity.csv") #read the file
activity$interval = as.factor(activity$interval) #intervals should be factors for futhers tasks
activity$date = as.Date(activity$date) #dates should be seen as dates for further tasks#
```


## What is mean total number of steps taken per day?
The next step is to find out total number of steps for ech day, the mean and th median for all days.
It was supposed to make a histogram for total number of steps, but I also made a plot for number of steps for each day.
```{r, echo = TRUE}
dates = unique(activity$date)
sums = vector()
for(i in 1:length(dates))
{sums[i]=sum(activity$steps[which(activity$date==dates[i] &!is.na(activity$steps))])}
sbd = data.frame(dates, sums) #total nmber of steps taken per day


hist(sbd$sums, breaks = 100) #histogram for the total number of steps taken each day
plot(sbd$dates, sbd$sums, type = "p") #showing how total number of steps distributes across all dates
mean(sbd$sums) #mean for the total number of steps taken each day
median(sbd$sums) #median for the total number of steps taken each day
```

So here you got it:
the median is 10395 step,
the mean is 9354.23

## What is the average daily activity pattern?
Talking about the day patterns, you need to do some job one avareging te thing across each interval.
Then I decided to create a data frame with names of intervals and its avaregaes.
There you can see the plot for that avarege and the calculus for mas avarege interval.
```{r}
intmeans = vector()
for(i in 1:length(intervals))
{intmeans[i]=mean(activity$steps[which(activity$interval==intervals[i] &!is.na(activity$steps))])}
meanbi = data.frame(intervals, intmeans)
plot(meanbi, type = "l", col = "green") #plot for the mean steps for each interval across all days
meanbi[which(meanbi$intmeans==max(meanbi$intmeans)),1] #5-minute interval that on average contains most number of steps 
```

## Imputing missing values
#3.1 Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
Going point by point: the total nuber of missing values: 2034
```{r}
length(which(is.na(activity$steps)==TRUE))
```

#3.2 NAs missing strategy (takign the decision)
Then need to tae a look at how those NAs are distributed.
I found out that there are several totally missed days, so the best alternative is to take the avarege for each interval.
```{r}
nas = activity[which(is.na(activity$steps)==TRUE),]
table(nas$date)
```

#3.3 filling NAs with mean for the interval, as filling it with mean fort he day is useless
Now I fill the missing values with that avarege, creating new data set and new variable.

```{r}
activity_na = activity
activity_na[,4] = rep(meanbi$intmeans, 61)

for(i in 1:length(activity_na$steps) )
{
  if (i %in% which(is.na(activity_na$steps))== TRUE) 
  {activity_na[i,5] <- activity_na[i,4]} 
  else
  {activity_na[i,5] <- activity_na[i,1]}
}
names(activity_na)[4] = "mean.for.the.interval"
names(activity_na)[5] = "value.after.NA.strategy"
```

#3.4 hist, mean, median with NAs filled in
Calculating all same thing for the new data set, with missing values being filled in
```{r}
dates = unique(activity$date)
nasums = vector()
for(i in 1:length(dates))
{nasums[i]=sum(activity_na$value.after.NA.strategy[which(activity$date==dates[i])])}
nasbd = data.frame(dates, nasums)
hist(nasbd$nasums)
mean(nasbd$nasums)
median(nasbd$nasums)
```
New median:10766.19
New mean: 10766.19
Somehow they are same, I do't know how.
I even checked with this:
```{r, echo = TRUE}
summary(nasbd$nasums)
```

#showing the difference
I also decided to show how exactly filled NAs changed the picture.
```{r}
plot(activity_na$date, activity_na$value.after.NA.strategy, col = "red", type = "l")
lines(activity$date, activity$steps, col = "green")
```
The red lines here are the missed data days filled with mean interval each. 
Could have filled with just avarege date figures, but then it would not be possible to see deeper than just a day.

## Are there differences in activity patterns between weekdays and weekends?
#4.1 defining weekdays and weekends
Taking the new set (activity_na) and filling in the "weekend/weekday" data.
```{r}
activity_na[,6] = weekdays(as.Date(activity$date))
days = unique(activity_na$V6)
for(i in 1:length(activity_na$steps))
{
  if(activity_na$V6[i] %in% days[1:5])
  {activity_na$V6[i] = "weekday"}
  #if(activity_na$V6[i] %in% days[6:7])
  else  {activity_na$V6[i]="weekend"}
}
names(activity_na)[6]="type.of.day"
```

#4.2 plotting weekends and weekdays steps
Now create two new data frames, one for weekend one for weekday avarege for each interval.
```{r}
weekdaymeans = vector()
for(i in 1:length(intervals))
{weekdaymeans[i]=mean(activity_na$value.after.NA.strategy[which(activity_na$type.of.day=="weekday")])}
daymeanbi = data.frame(intervals, weekdaymeans)

weekendmeans = vector()
for(i in 1:length(intervals))
{weekendmeans[i]=mean(activity_na$value.after.NA.strategy[which(activity_na$type.of.day=="weekend")])}
endmeanbi = data.frame(intervals, weekendmeans)

plot(endmeanbi$weekendmeans, col = "red", type = "l")
lines(daymeanbi$weekdaymeans, col = "green")
temp <- legend("topright", legend = c("Weekday ", "Weekend"), text.width = strwidth("1,000,000"),lty = 1, col = c("green", "red"), xjust = 1, yjust = 1, title = "Line Types")
```

It can be seen from the graph, that weekend activity (the red line) starts later and ends later than weekday (the green line), and is more intense in the day time.
Overall, the sums of steps for the avarege weekend and weekday are here:
```{r}
sum(endmeanbi$weekendmeans)
sum(daymeanbi$weekdaymeans)
```
Weekend: 12201.85
Weekday: 10255.85

So yes, we can say that on avarege fitnes tracker users are kore active on weekend then on weekdays.
